---
title: "Untitled"
output: html_document
---

```{r}
# Load the relevant package
library(tidyverse)
library(stats)
library(GGally)

# Load the data
bustabit <- read.csv("bustabit.csv")

# Find the highest multiplier (BustedAt value) achieved in a game
bustabit %>%
  arrange(desc(BustedAt)) %>%
  slice(1)
```

2. Deriving relevant features for clustering
```{r}
# Create the new feature variables
bustabit_features <- bustabit %>%
  mutate(CashedOut = ifelse(is.na(CashedOut), BustedAt + 0.01, CashedOut),
         Profit = ifelse(is.na(Profit), 0, Profit),
         Losses = ifelse(Profit==0, -1*Bet, 0),
         GameWon = ifelse(Profit==0, 0, 1),
         GameLost = ifelse(Profit==0, 1, 0))

# Look at the first five rows if the feature data
head(bustabit_features, 5)
```

3. Creating per-player statistics
```{r}
# Group by players to create per-player summary statistics
bustabit_clus <- bustabit_features %>%
  group_by(Username) %>%
  summarize(AverageCashedOut = mean(CashedOut), 
            AverageBet = mean(Bet),
            TotalProfit = sum(Profit),
            TotalLosses = sum(Losses), 
            GamesWon = sum(GameWon),
            GamesLost = sum(GameLost))

# View the first five rows of the data
head(bustabit_clus, n = 5)
```

4. Scaling and normalization of the derived features
```{r}
# Create the mean-sd standardization function
mean_sd_standard <- function(x) {
  (x-mean(x)) / sd(x)
}

# Apply the function to each numeric variables in the clustering
bustabit_standardized <- bustabit_clus %>%
  mutate_if(is.numeric, mean_sd_standard)

# Summarize our standardized data
summary(bustabit_standardized)
```

5. Cluster the player data using K means
```{r}
# Set the seed
set.seed(100)

# Cluster the players using kmeans with five clusters
cluster_solution <- kmeans(bustabit_standardized[,-1], centers = 5)

# Store the cluster assignments back into the clustering data frame object
bustabit_clus$cluster <- factor(cluster_solution$cluster)

# Look at the distribution of cluster assignments
table(bustabit_clus$cluster)
```

6. Compute averages for each cluster
```{r}
# Group by the cluster assignment and calculate averages
bustabit_clus_avg <- bustabit_clus %>%
  group_by(cluster) %>%
  summarize_if(is.numeric, mean)

# View the resulting table
bustabit_clus_avg
```

7. Visualize the clusters with a Parallel Coordinate Plot
```{r}
# Create the min-max scaling function
min_max_standard <- function(x) {
  (x-min(x)) / (max(x) - min(x))
}

# Apply this function to each numeric variable in the bustabit_clus_avg object
bustabit_avg_minmax <- bustabit_clus_avg %>%
  mutate_if(is.numeric, min_max_standard)

# Create a parallel coordinate plot of the values
ggparcoord(bustabit_avg_minmax, columns = 2:ncol(bustabit_avg_minmax),
           groupColumn = "cluster", scale = "globalminmax", order = "skewness")
```

8. Visualize the clusters with Principal Components
```{r}
# Calculate the principal components of the standardized data
my_pc <- as.data.frame(prcomp(bustabit_standardized[,-1])$x)

# Store the cluster assignments in the new data frame
my_pc$cluster <- bustabit_clus$cluster

# Use the ggplot() to plot PC2 vs PC1, and color by the cluster assignment
p1 <- ggplot(data = my_pc, aes(x = PC1, y = PC2, color = cluster)) + geom_point()

p1
```

9. Analyzing the groups of gamblers our solution uncovered
```{r}
# Assign cluster names to clusters 1 through 5 in order
cluster_names <- c(
    "Risky Commoners",
    "High Rollers",
    "Risk Takers",
    "Cautious Commoners",
    "Strategic Addicts"
)

# Append the cluster names to the cluster means table
bustabit_clus_avg_names <- bustabit_clus_avg %>%
  cbind(Name = cluster_names)

# View the cluster means table with your appended cluster names
bustabit_clus_avg_names
```

